Treatment Outcome Visualization

### 2023-06-18 Diagnosis_Vis.Rmd
Pre/Post Plots

@@

title = "Baseline Depression for Patients By Disorder Type"
subtitle = "Upon entrence of the partial hospitalization program, scores of anxiety are reported from the _Clinically Useful Depression Outcome Scale_ (CUDOS). The CUXOS is a 18 item measure used to assess different symptoms of depression; A cut off score of 20 suggests a positive screen for probable major depression. Mean CUDOS scores are reported for each disorder type below (i.e., _the black line_)."
cap = "**Symptom severity:** 0-10 (non-depression), 11-20 (minimal depression), 21-30 (mild anxiety), 31-45 (moderate depression), 41+ (severe depression) <br><br> **Source:**: Methods to Improve Diagnostic Assessment and Services (MIDAS) project"

>>>>

### Export as CSV
write.csv(Mothership,"~/Desktop/Coding/data/Mothership_Diag.csv")

@@

title = "Baseline Depression for Patients By Disorder Type"
subtitle = "Upon entrence of the partial hospitalization program, scores of anxiety are reported from the _Clinically Useful Depression Outcome Scale_ (CUDOS). The CUXOS is a 18 item measure used to assess different symptoms of depression; A cut off score of 20 suggests a positive screen for probable major depression. Mean CUDOS scores are reported for each disorder type below (i.e., _the black line_)."
cap = "**Symptom severity:** 0-10 (non-depression), 11-20 (minimal depression), 21-30 (mild depression), 31-45 (moderate depression), 46+ (severe depression) <br><br> **Source:**: Methods to Improve Diagnostic Assessment and Services (MIDAS) project"

### 2023-06-18 Treatment_Outcomes.Rmd
Pre/Post Plots

#Data
Mothership<-read.csv("~/Desktop/Coding/data/Mothership_Vis.csv")
Mothership<-mutate(Mothership,
      DC_status_1 = as.factor(DC_status_1),
      Education_1 = as.factor(Education_1),
      Gender_1 = as.factor(Gender_1),
      Relationship_1 = as.factor(Relationship_1),
      Sexuality_1 = as.factor(Sexuality_1),
      IPT_track_1 = as.factor(IPT_track_1),
      Race_1 = as.factor(Race_1),
      Education_1 = as.factor(Education_1),
      Duration = as.factor(Duration),
      )

@@

@@

```{r, include=FALSE}
# Depression symptoms df
# CUDOS_Score <- select(Mothership,
#                       Days_complete_1,Day1_CUDOS:Day40_CUDOS)
# 
# # Anxiety symptoms df
# CUXOS_Score <- select(Mothership,
#                       Days_complete_1,Day1_CUXOS:Day40_CUXOS)

```{r, include=FALSE}
#1. First adjust the their values for length of stay above 40 to 40, because the measures don't go above 40 days.
# 
# 
# Mothership_post <- Mothership %>%
#   mutate(Days_complete_W = ifelse(Days_complete_1 > 40, 40,Days_complete_1))
# 
# 
# 
# Mothership <- Mothership %>%
#   mutate(Days_complete_W = ifelse(Days_complete_1 > 40, 40,Days_complete_1))
#          
# 
# Mothership$Day1_CUDOS
# Mothership$Days_complete_1
# Mothership$Days_complete_W
# 
# view(Mothership)
# #2. Then create df with CUDOS and CUXOS scores
#   # Depression
# 
#     CUDOS_Score <- select(Mothership,
#                       Day1_CUDOS:Day40_CUDOS)
# 
#   # Anxiety
#     CUXOS_Score <- select(Mothership,
#                       Day1_CUXOS:Day40_CUXOS)
# 
# 
# #3. Extract the valeues for each patient at their last day (length of stay)
#     ##### Why is it not includeing NAs.
#     
# 
# CUDOS_post = CUDOS_Score[cbind(seq_along(Mothership$Days_complete_W),Mothership$Days_complete_W)]
# CUXOS_post = CUXOS_Score[cbind(1:length(Mothership$Days_complete_W),Mothership$Days_complete_W)]
# 
# seq_along(Mothership$Days_complete_W)
# Mothership$Days_complete_W
# CUDOS_post
# 
# Mothership %>%
#   mutate(Dep_post = CUDOS_post)
# 
# 
# Mothership %>%
#   mutate(Dep_post = CUDOS_post)
# 
# CUDOS_post
# #4. Extract the value for their first day
# CUDOS_pre<-Mothership$Day1_CUDOS
# CUXOS_pre<-Mothership$Day1_CUXOS
# AnxDep_PrePost_test <- data.frame(cbind(Mothership$ID1,CUDOS_pre,CUDOS_post,CUXOS_pre,CUXOS_post))
# AnxDep_PrePost_test
# colnames(AnxDep_PrePost)[1] = "ID1"
# 
# AnxDep_PrePost_test
# ## This is where I am Figure out where to go from here *******************
# #5. Create long dataframe
# 
# AnxDep_PrePost_Lon <- gather(AnxDep_PrePost,ID1:CUXOS_post,factor_key = "T")
# 
# AnxDep_PrePost_Long <- pivot_longer(AnxDep_PrePost,ID1:CUXOS_post)
# 
# 
# AnxDep_PrePost_Long
# ?pivot_longer()
# 
# AnxDep_PrePost_Long <- AnxDep_PrePost_Long[order(AnxDep_PrePost_Long$ID1),]
# 
# Mothership_Long_RDQ <- gather(Mothership_RDQ,rm,RDQ,rdqPRE_coping:rdqPOST_sym,
#                               factor_key = "T")
# Mothership_Long_RDQ

```{r, include=FALSE}
# Mothership_Post <- Mothership %>%
#   filter(!is.na(Days_complete_1)) %>%
#   mutate(Days_complete_W = ifelse(Days_complete_1 > 40, 40,Days_complete_1)) 
#   
# 
# Mothership_Post$Days_complete_W
# 1:length(Mothership_Post$Days_complete_W)
# 
# # Depression
# CUDOS_Score <- select(Mothership_Post, Day1_CUDOS:Day40_CUDOS)
# 
# CUDOS_Score
# 
# 1:length(Mothership_Post$Days_complete_W)
# 
# # Anxiety
# CUXOS_Score <- select(Mothership_Post, Day1_CUXOS:Day40_CUXOS)
# 
# values <- cbind(seq_along(Mothership_Post$Days_complete_W),Mothership_Post$Days_complete_W[1:8610])
# values
# CUDOS_post = CUDOS_Score[cbind(seq_along(Mothership_Post$Days_complete_W)[1:8610],Mothership_Post$Days_complete_W[1:8610])]
# CUDOS_post
# 
# 
# CUDOS_post = CUDOS_Score[cbind(1:length(Mothership_Post$Days_complete_W),Mothership_Post$Days_complete_W)]
# CUXOS_post = CUXOS_Score[cbind(1:length(Mothership_Post$Days_complete_W),Mothership_Post$Days_complete_W)]
# 
# CUDOS_post
# 
# values <- cbind(1:length(Mothership_Post$Days_complete_W),Mothership_Post$Days_complete_W)
# values
# Mothership_Post<- Mothership_Post %>%
#   mutate(Dep_post = CUDOS_post)
```

>>>>

#Data
Mothership<-read.csv("~/Desktop/Coding/data/Mothership_Diag.csv")
Mothership<-mutate(Mothership,
      DC_status_1 = as.factor(DC_status_1),
      Education_1 = as.factor(Education_1),
      Gender_1 = as.factor(Gender_1),
      Relationship_1 = as.factor(Relationship_1),
      Sexuality_1 = as.factor(Sexuality_1),
      IPT_track_1 = as.factor(IPT_track_1),
      Race_1 = as.factor(Race_1),
      Education_1 = as.factor(Education_1),
      Duration = as.factor(Duration),
      )

@@

Mothership_Mean_RDQ

@@

```{r, include=FALSE}
# 1. First adjust the their values for length of stay above 40 to 40, because the measures don't go above 40 days.

Mothership_post <- Mothership %>%
  mutate(Days_complete_W = ifelse(Days_complete_1 > 40, 40,Days_complete_1),

#2. Create pre_scores
      CUXOS_pre = Day1_CUXOS,
      CUDOX_pre = Day1_CUDOS
      )

#3. Then create df with CUDOS and CUXOS scores
  # Depression

    CUDOS_Score <- select(Mothership,
                      Day1_CUDOS:Day40_CUDOS)
    CUDOS_Score <- data.frame(CUDOS_Score)

  # Anxiety
    CUXOS_Score <- select(Mothership,
                      Day1_CUXOS:Day40_CUXOS)
    CUXOS_Score <- data.frame(CUDOS_Score)

    view(Mothership)

#4. Extract the valeues for each patient at their last day (length of stay), and create a CUDOS and CUXOS post score.

Mothership <- Mothership %>%
  mutate(
    CUDOS_post = CUDOS_Score[cbind(seq_along(1:8667),CUDOS_Score$Days_complete_W)],
    CUXOS_post = CUXOS_Score[cbind(seq_along(1:8667),CUDOS_Score$Days_complete_W)]
  )

#5. create new DF with Varaibles we want

AnxDep_PrePost<- data.frame(cbind(Mothership$ID1,Mothership$Disorder_Type,CUDOS_pre,CUDOS_post,CUXOS_pre,CUXOS_post)) %>%
  filter(!is.na(Mothership$Disorder_Type))
colnames(AnxDep_PrePost)[1:2] = c("ID1","Disorder_Type")

#6. Create long data frame

AnxDep_PrePost_Long <- AnxDep_PrePost %>% 
  pivot_longer(
    cols = CUDOS_pre:CUXOS_post, 
    names_to = "Test",
    values_to = "Score"
  )
AnxDep_PrePost_Long <-AnxDep_PrePost_Long %>%
  mutate(
    PrePost = ifelse(
      Test == "CUDOS_pre", "Pre",
      ifelse(Test == "CUXOS_pre" , "Pre", "Post"
    )),
    Score = as.integer(Score),
    Test = as.factor(Test),
    Test = as.factor(ifelse(Test == "CUDOS_pre",  "CUDOS",
                     ifelse(Test == "CUDOS_post", "CUDOS", "CUXOS"
                     ))))

#7. Convert to data frame with means of values to be plotted
AnxDep_PrePost_mean<-AnxDep_PrePost_Long %>%
  group_by(Test,PrePost) %>%
  summarise(Score=mean(Score,na.rm=TRUE))

#8. Pre Post CUDOS/CUXOS Plot
ggplot(data=AnxDep_PrePost_mean,
       aes(x=factor(PrePost,level=c("Pre","Post")),y=Score,
           group=Test,color=Test))+
  geom_point(size=2.5)+
  geom_line(group=AnxDep_PrePost_mean$PrePost,size=1.5)+
  scale_x_discrete(name="Time Point")+
  scale_y_continuous(name="CUDOS/CUXOS Score")+
  ggtitle("Mean Pre and Post Scores of the Anxiety and Depression")+
  theme(panel.grid.major.y = element_line(color = "grey",size = 0.5,linetype = 2))+
  theme(panel.background=NULL)+
  theme(legend.title = element_blank())
 # scale_color_manual(values=c("1"="#8956BB","2"="#4C4452","3"="#B1A8B9",
                             # "4"="#00BCA3","5"="#008570"),
                     #labels=c("Coping Skills","Positive Mental Health",
                             # "Functioning","Well-Being","Symptoms"))

```
Anxiety / Depression Pre_Post by disorder - line plot
```{r}
#1. Creaate new datasets with only anxiety / depresssion

ANX_PrePost<- AnxDep_PrePost_Long %>%
  filter(Test == "CUXOS")

DEP_PrePost<- AnxDep_PrePost_Long %>%
  filter(Test == "CUDOS")

#2. Create DF grouped buy disorder type
ANX_PrePost_mean<-ANX_PrePost %>%
  group_by(Disorder_Type,PrePost) %>%
  summarise(Score=mean(Score,na.rm=TRUE))

DEP_PrePost_mean<-DEP_PrePost %>%
  group_by(Disorder_Type,PrePost) %>%
  summarise(Score=mean(Score,na.rm=TRUE))

#3. Plot em

#### ANX
ggplot(data=ANX_PrePost_mean,
       aes(x=factor(PrePost,level=c("Pre","Post")),y=Score,
           group=Disorder_Type,color=Disorder_Type))+
  geom_point(size=2.5)+
  geom_line(size=1)+
  scale_x_discrete(name="Time Point")+
  scale_y_continuous(name="CUXOS Score")+
  labs(title = "Mean Pre and Post Anxiety Scores By Disorder Type",
       subtitle = "Self Report anxiety was assessed using the CUXOS",
       caption = "Symptom severity: 0-10 (remission), 11-20 (minimal), 21-30 (mild), 31-40 (moderate), 41+ (severe)")+
  theme(panel.grid.major.y = element_line(color = "grey",size = 0.5,linetype = 2))+
  theme(panel.background=NULL)+
  theme(legend.title = element_blank())

#### DEP
ggplot(data=DEP_PrePost_mean,
       aes(x=factor(PrePost,level=c("Pre","Post")),y=Score,
           group=Disorder_Type,color=Disorder_Type))+
  geom_point(size=2.5)+
  geom_line(size=1)+
  scale_x_discrete(name="Time Point")+
  scale_y_continuous(name="CUDOS Score")+
  labs(title = "Mean Pre and Post Depression Scores By Disorder Type",
       subtitle = "Self report depression was assessed using the CUDOS",
       caption = "Symptom severity: 0-10 (remission), 11-20 (minimal), 21-30 (mild), 31-45 (moderate), 46+ (severe)")+
  theme(panel.grid.major.y = element_line(color = "grey",size = 0.5,linetype = 2))+
  theme(panel.background=NULL)+
  theme(legend.title = element_blank())
```

Depression/ Anxiety pre/ post Dumbell plots
```{r}
Pre_ANX<-ANX_PrePost_mean %>%
  filter(PrePost == "Pre")

Post_ANX<-ANX_PrePost_mean %>%
  filter(PrePost == "Post")

Plot_Text<- dplyr::count(ANX_PrePost, Disorder_Type)
Plot_Text

ggplot(data=Pre_ANX,
       aes(x=Disorder_Type,y=Score,
           group=Disorder_Type,color=Disorder_Type))+
  geom_segment(aes(x=Disorder_Type, xend=Post_ANX$Disorder_Type, y=Score, yend=Post_ANX$Score),color = "#aeb6bf",
              size = 4, #Note that I sized the segment to fit the points
              alpha = .5) +
  geom_point( aes(x=Disorder_Type, y=Score), color= "#F90000", size=3.5) +
  geom_point( aes(x=Post_ANX$Disorder_Type, y=Post_ANX$Score), color= "#0CC541", size=3.5) +
  labs(title = "Mean Pre and Post Anxiety Scores By Disorder Type",
       subtitle = "Self Report anxiety was assessed using the CUXOS",
       caption = "Symptom severity: 0-10 (remission), 11-20 (minimal), 21-30 (mild), 31-40 (moderate), 41+ (severe)")+
  scale_y_reverse()+
  # limits = c(10, 50))+
  coord_flip() +
  theme_classic()+
  theme(#legend.position = "none",
         plot.background = element_rect(
      fill = "white",
      colour = "white"))

sum(as.factor(ANX_PrePost$Disorder_Type))

Pre_DEP<-DEP_PrePost_mean %>%
  filter(PrePost == "Pre")

Post_DEP<-DEP_PrePost_mean %>%
  filter(PrePost == "Post")

ggplot(data=Pre_DEP,
       aes(x=Disorder_Type,y=Score,
           group=Disorder_Type,color=Disorder_Type))+
  geom_segment(aes(x=Disorder_Type, xend=Post_DEP$Disorder_Type, y=Score, yend=Post_DEP$Score),color = "#aeb6bf",
              size = 4, #Note that I sized the segment to fit the points
              alpha = .5) +
  geom_point( aes(x=Disorder_Type, y=Score), color= "#F90000", size=3.5) +
  geom_point( aes(x=Post_DEP$Disorder_Type, y=Post_DEP$Score), color= "#0CC541", size=3.5) +
  labs(title = "Mean Pre and Post Depression Scores By Disorder Type",
       subtitle = "Self Report anxiety was assessed using the CUDOS",
       caption = "Symptom severity: 0-10 (remission), 11-20 (minimal), 21-30 (mild), 31-45 (moderate), 46+ (severe)")+
  scale_y_reverse()+
  # limits = c(10, 50))+
  coord_flip() +
  theme_classic()+
  theme(#legend.position = "none",
         plot.background = element_rect(
      fill = "white",
      colour = "white"))
```
